---
title: "CH and HIV analysis"
author: "Christelle Colin-Leitzinger"
date: '`r Sys.Date()`'
output: 
  html_document:
    toc: true
    toc_float: true
    toc_collapsed: false
    theme: united
    highlight: pygments
    df_print: kable
editor_options: 
  chunk_output_type: console
---

<style type="text/css">

.figure {
    margin-top: 100px;
    margin-bottom: 100px;
}

table {
    margin-top: 10px;
    margin-bottom: 25px !important;
}

th, td { padding: 5px; }

</style>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      warning = FALSE,
                      message = FALSE,
                      cache = FALSE,
                      fig.align='center'
                      )
```

```{r library}
library(tidyverse)
library(gtsummary)
library(survival)
library(survminer)
theme_set(theme_classic())
theme_gtsummary_compact()
```

```{r}
CH_HIV_data <- 
  read_rds(paste0(here::here(), "/CH_HIV_data.rds")) %>% 
  group_by(MRN) %>% 
  mutate(n = n(), .before = MRN) %>% 
  ungroup() %>% 
  filter(baseline_sample == 1 |
           (n == 1 & baseline_sample == 2)) %>% 
  select(-n)
```

# Table 1
```{r table 1}
CH_HIV_data %>% 
  select(age_at_diagnosis_corrected, age_at_specimen,
         hiv_status, CH_status,
         starts_with("grim"), starts_with("tnsc"),
         starts_with("horvath"),
         starts_with("pheno_"),
         starts_with("dunedin"),
         new_race, new_ethnicity, sex, 
         new_tnm_stage, clinical_stage, 
         vital_status, tx_summary, treatment_status,
         specimen_to_surgery_lag,
         specimen_to_treatment_lag, specimen_tx_lag,
         new_primary_site
         ) %>% 
  tbl_summary() %>% 
  bold_labels()
```

```{r plot clocks, fig.height=14, fig.width=12}
CH_HIV_data %>% 
  select(CH_status,
         starts_with("grim"), starts_with("tnsc"),
         starts_with("horvath"),
         starts_with("pheno_"),
         starts_with("dunedin")) %>% 
  filter(!is.na(CH_status)) %>% 
  pivot_longer(cols = -CH_status) %>% 
  ggplot(aes(x= CH_status, y= value, color= CH_status))+
  geom_boxplot()+
  scale_color_manual(values = c("blue", "red"))+
  facet_wrap(.~ name, scales = "free", ncol = 3)+
  theme(axis.title.y = element_blank())+
  stat_compare_means()

CH_HIV_data %>% 
  select(hiv_status,
         starts_with("grim"), starts_with("tnsc"),
         starts_with("horvath"),
         starts_with("pheno_"),
         starts_with("dunedin")) %>% 
  filter(!is.na(hiv_status)) %>% 
  pivot_longer(cols = -hiv_status) %>% 
  ggplot(aes(x= hiv_status, y= value, color= hiv_status))+
  geom_boxplot()+
  scale_color_manual(values = c("darkgrey", "purple"))+
  facet_wrap(.~ name, scales = "free", ncol = 3)+
  theme(axis.title.y = element_blank())+
  stat_compare_means()
```

# Table 2 - stratified by CH
```{r table 2}
CH_HIV_data %>% 
  select(age_at_diagnosis_corrected, age_at_specimen,
         hiv_status, CH_status,
         starts_with("grim"), starts_with("tnsc"),
         starts_with("horvath"),
         starts_with("pheno_"),
         starts_with("dunedin"),
         new_race, new_ethnicity, sex, 
         new_tnm_stage, clinical_stage, 
         vital_status, tx_summary, treatment_status,
         specimen_to_surgery_lag,
         specimen_to_treatment_lag, specimen_tx_lag,
         new_primary_site
         ) %>% 
  tbl_summary(by = CH_status) %>% 
  bold_labels() %>% 
  add_p() %>% bold_p(t=.05) %>% add_overall()
```

```{r}
tbl1 <- CH_HIV_data %>% 
  filter(CH_status == "CH") %>% 
  select(hiv_status,
         new_primary_site
         ) %>% 
  tbl_summary(by = hiv_status) %>% 
  bold_labels() %>% 
  add_p() %>% bold_p(t=.05) %>% add_overall()
tbl2 <- CH_HIV_data %>% 
  filter(CH_status == "NO") %>% 
  select(hiv_status,
         new_primary_site
         ) %>% 
  tbl_summary(by = hiv_status) %>% 
  bold_labels() %>% 
  add_p() %>% bold_p(t=.05) %>% add_overall()
tbl_merge(list(tbl1, tbl2), tab_spanner = c("CH", "No CH"))
```

# Table 3 - stratified by HIV
```{r}
CH_HIV_data %>% 
  select(age_at_diagnosis_corrected, age_at_specimen,
         hiv_status, CH_status,
         starts_with("grim"), starts_with("tnsc"),
         starts_with("horvath"),
         starts_with("pheno_"),
         starts_with("dunedin"),
         new_race, new_ethnicity, sex, 
         new_tnm_stage, clinical_stage, 
         vital_status, tx_summary, treatment_status,
         specimen_to_surgery_lag,
         specimen_to_treatment_lag, specimen_tx_lag,
         new_primary_site
         ) %>% 
  tbl_summary(by = hiv_status) %>% 
  bold_labels() %>% 
  add_p() %>% bold_p(t=.05) %>% add_overall()
```

<!-- # Table 4 - stratified by CH and HIV -->


# Survival analysis
**Note : Survival time is calculated from date of treatment to date of death**  
<div class = "row">
<div class = "col-md-6">
```{r Survival, fig.height = 7}
ggsurvplot(survfit(Surv(month_at_os, os_event) ~ CH_status, 
                   data=CH_HIV_data),
           title = "OS Analysis",
           font.main = c(20, "bold", "black"),
           font.x = c(18, "bold", "black"),
           font.y = c(18, "bold", "black"),
           font.legend = c(16, "black"),
           font.tickslab = c(16, "bold", "black"),
           size = 1.5,

           xlab = "Time in months",
           legend = "top",
           legend.title = "",
           pval = TRUE,
           conf.int = FALSE,
           # Censor
           censor = TRUE
) + guides(colour = guide_legend(ncol = 1))
```
</div>

<div class = "col-md-6">
```{r, fig.height = 7}
ggsurvplot(survfit(Surv(month_at_os, os_event) ~ hiv_status, 
                   data=CH_HIV_data),
           title = "OS Analysis",
           font.main = c(20, "bold", "black"),
           font.x = c(18, "bold", "black"),
           font.y = c(18, "bold", "black"),
           font.legend = c(16, "black"),
           font.tickslab = c(16, "bold", "black"),
           size = 1.5,

           xlab = "Time in months",
           legend = "top",
           legend.title = "",
           pval = TRUE,
           conf.int = FALSE,
           # Censor
           censor = TRUE
) + guides(colour = guide_legend(ncol = 1))
```
</div>
</div>

```{r cox}
tbl1 <- CH_HIV_data %>% 
  select(CH_status, 
         age_at_diagnosis_corrected,
         new_race, new_ethnicity, sex, hiv_status, clinical_stage) %>% 
  tbl_uvregression(method = survival::coxph, 
                   y = (Surv(time = CH_HIV_data$month_at_os, 
                             event = CH_HIV_data$os_event)),
                   exponentiate = TRUE) %>% 
  bold_p(t = .05) %>% add_nevent(location = "level") %>% add_n(location = "level") %>%
  bold_labels() %>% italicize_levels()
tbl2 <- coxph(Surv(time = CH_HIV_data$month_at_os, 
             event = CH_HIV_data$os_event) ~ 
               CH_status + age_at_diagnosis_corrected + new_race + new_ethnicity + sex + hiv_status + clinical_stage, 
             data =  CH_HIV_data) %>%
  tbl_regression(exponentiate = TRUE) %>% bold_p(t = .05)

tbl_merge(list(tbl1, tbl2), tab_spanner = c("**Univariable**", "**Multivariable**"))

tbl1 <- CH_HIV_data %>% 
  select(CH_status, 
         hiv_status) %>% 
  tbl_uvregression(method = survival::coxph, 
                   y = (Surv(time = CH_HIV_data$month_at_os, 
                             event = CH_HIV_data$os_event)),
                   exponentiate = TRUE) %>% 
  bold_p(t = .05) %>% add_nevent(location = "level") %>% add_n(location = "level") %>%
  bold_labels() %>% italicize_levels()
tbl2 <- coxph(Surv(time = CH_HIV_data$month_at_os, 
             event = CH_HIV_data$os_event) ~ 
               CH_status + hiv_status, 
             data =  CH_HIV_data) %>%
  tbl_regression(exponentiate = TRUE) %>% bold_p(t = .05)

tbl_merge(list(tbl1, tbl2), tab_spanner = c("**Univariable**", "**Multivariable**"))
```
Cox Proportion Hazard Assumption - No violation of the model
```{r}
coxph(Surv(time = CH_HIV_data$month_at_os, 
             event = CH_HIV_data$os_event) ~ 
               CH_status + age_at_diagnosis_corrected + new_race + new_ethnicity + sex + hiv_status + clinical_stage, 
             data =  CH_HIV_data) %>% 
  cox.zph()
```

```{r}
# tbl1 <- CH_HIV_data %>% 
#   select(CH_status, 
#          age_at_diagnosis_corrected,
#          new_race, new_ethnicity, sex, hiv_status, clinical_stage) %>% 
#   tbl_uvregression(method = survival::coxph, 
#                    y = (Surv(time = CH_HIV_data$month_at_os, 
#                              event = CH_HIV_data$os_event)),
#                    exponentiate = TRUE) %>% 
#   bold_p(t = .05) %>% add_nevent(location = "level") %>% add_n(location = "level") %>%
#   bold_labels() %>% italicize_levels()
# tbl2 <- coxph(Surv(time = CH_HIV_data$month_at_os, 
#              event = CH_HIV_data$os_event) ~ 
#                CH_status + age_at_diagnosis_corrected + new_race + new_ethnicity + sex + hiv_status + clinical_stage, 
#              data =  CH_HIV_data) %>%
#   tbl_regression(exponentiate = TRUE) %>% bold_p(t = .05)
# 
# tbl_merge(list(tbl1, tbl2), tab_spanner = c("**Univariable**", "**Multivariable**"))
```
<br>
<br>

# Startified Survival by HIV/CH
```{r}
ggsurvplot(survfit(Surv(month_at_os, os_event) ~ hiv_status + CH_status, 
                   data=CH_HIV_data),
           title = "OS Analysis",
           font.main = c(20, "bold", "black"),
           font.x = c(18, "bold", "black"),
           font.y = c(18, "bold", "black"),
           font.legend = c(16, "black"),
           font.tickslab = c(16, "bold", "black"),
           size = 1.5,

           xlab = "Time in months",
           legend = "top",
           legend.title = "",
           color = "hiv_status",
           linetype = "CH_status",
           pval = TRUE,
           conf.int = FALSE,
           # Censor
           censor = TRUE
) + guides(colour = guide_legend(ncol = 1),
           linetype = guide_legend(ncol = 1))
```

## In HIV + population
```{r}
HIV_pos <- CH_HIV_data %>% 
  filter(hiv_status == "positive")
tbl1 <- HIV_pos %>% 
  select(CH_status, 
         age_at_diagnosis_corrected,
         new_race, new_ethnicity, sex, clinical_stage) %>% 
  tbl_uvregression(method = survival::coxph, 
                   y = (Surv(time = HIV_pos$month_at_os, 
                             event = HIV_pos$os_event)),
                   exponentiate = TRUE) %>% 
  bold_p(t = .05) %>% add_nevent(location = "level") %>% add_n(location = "level") %>%
  bold_labels() %>% italicize_levels()
tbl2 <- coxph(Surv(time = HIV_pos$month_at_os, 
             event = HIV_pos$os_event) ~ 
               CH_status + age_at_diagnosis_corrected + new_race + new_ethnicity + sex + clinical_stage, 
             data =  HIV_pos) %>%
  tbl_regression(exponentiate = TRUE) %>% bold_p(t = .05)

tbl_merge(list(tbl1, tbl2), tab_spanner = c("**Univariable**", "**Multivariable**"))
```

## In HIV - population
```{r}
HIV_neg <- CH_HIV_data %>% 
  filter(hiv_status == "negative")
tbl1 <- HIV_neg %>% 
  select(CH_status, 
         age_at_diagnosis_corrected,
         new_race, new_ethnicity, sex, clinical_stage) %>% 
  tbl_uvregression(method = survival::coxph, 
                   y = (Surv(time = HIV_neg$month_at_os, 
                             event = HIV_neg$os_event)),
                   exponentiate = TRUE) %>% 
  bold_p(t = .05) %>% add_nevent(location = "level") %>% add_n(location = "level") %>%
  bold_labels() %>% italicize_levels()
tbl2 <- coxph(Surv(time = HIV_neg$month_at_os, 
             event = HIV_neg$os_event) ~ 
               CH_status + age_at_diagnosis_corrected + new_race + new_ethnicity + sex + clinical_stage, 
             data =  HIV_neg) %>%
  tbl_regression(exponentiate = TRUE) %>% bold_p(t = .05)

tbl_merge(list(tbl1, tbl2), tab_spanner = c("**Univariable**", "**Multivariable**"))
```

## In No CH population
```{r}
No_CH <- CH_HIV_data %>% 
  filter(CH_status == "NO")
tbl1 <- No_CH %>% 
  select(hiv_status, 
         age_at_diagnosis_corrected,
         new_race, new_ethnicity, sex, clinical_stage) %>% 
  tbl_uvregression(method = survival::coxph, 
                   y = (Surv(time = No_CH$month_at_os, 
                             event = No_CH$os_event)),
                   exponentiate = TRUE) %>% 
  bold_p(t = .05) %>% add_nevent(location = "level") %>% add_n(location = "level") %>%
  bold_labels() %>% italicize_levels()
tbl2 <- coxph(Surv(time = No_CH$month_at_os, 
             event = No_CH$os_event) ~ 
               hiv_status + age_at_diagnosis_corrected + new_race + new_ethnicity + sex + clinical_stage, 
             data =  No_CH) %>%
  tbl_regression(exponentiate = TRUE) %>% bold_p(t = .05)

tbl_merge(list(tbl1, tbl2), tab_spanner = c("**Univariable**", "**Multivariable**"))
```

## In CH population
```{r}
CH_data <- CH_HIV_data %>% 
  filter(CH_status == "CH")
tbl1 <- CH_data %>% 
  select(hiv_status, 
         age_at_diagnosis_corrected,
         new_race, new_ethnicity, sex, clinical_stage) %>% 
  tbl_uvregression(method = survival::coxph, 
                   y = (Surv(time = CH_data$month_at_os, 
                             event = CH_data$os_event)),
                   exponentiate = TRUE) %>% 
  bold_p(t = .05) %>% add_nevent(location = "level") %>% add_n(location = "level") %>%
  bold_labels() %>% italicize_levels()
tbl2 <- coxph(Surv(time = CH_data$month_at_os, 
             event = CH_data$os_event) ~ 
               hiv_status + age_at_diagnosis_corrected + new_race + new_ethnicity + sex + clinical_stage, 
             data =  CH_data) %>%
  tbl_regression(exponentiate = TRUE) %>% bold_p(t = .05)

tbl_merge(list(tbl1, tbl2), tab_spanner = c("**Univariable**", "**Multivariable**"))
```

```{r}
ggsurvplot(survfit(Surv(month_at_os, os_event) ~ CH_status, 
                   data=CH_HIV_data),
           title = "OS Analysis",
           font.main = c(20, "bold", "black"),
           font.x = c(18, "bold", "black"),
           font.y = c(18, "bold", "black"),
           font.legend = c(16, "black"),
           font.tickslab = c(16, "bold", "black"),
           size = 1.5,

           xlab = "Time in months",
           legend = "top",
           legend.title = "CH status",
           legend.labs = c("No", "CH"),
           palette = c("blue", "red"),
           pval = TRUE,
           conf.int = FALSE,
           # Censor
           censor = TRUE
) + guides(colour = guide_legend(ncol = 1))
```

```{r}
ggsurvplot(survfit(Surv(month_at_os, os_event) ~ hiv_status, 
                   data=CH_HIV_data),
           title = "OS Analysis",
           font.main = c(20, "bold", "black"),
           font.x = c(18, "bold", "black"),
           font.y = c(18, "bold", "black"),
           font.legend = c(16, "black"),
           font.tickslab = c(16, "bold", "black"),
           size = 1.5,

           xlab = "Time in months",
           legend = "top",
           legend.title = "HIV status",
           legend.labs = c("negative", "positive"),
           palette = c("darkgrey", "purple"),
           pval = TRUE,
           conf.int = FALSE,
           # Censor
           censor = TRUE
) + guides(colour = guide_legend(ncol = 1))
```

# Logistic regression
```{r}
glm(CH_status ~
      hiv_status + age_at_diagnosis_corrected + new_race + sex + treatment_status,
    family = "binomial",
    data = CH_HIV_data) %>%
  tbl_regression(exponentiate = TRUE,
                 intercept = TRUE) %>%
  modify_spanning_header(everything() ~ "CH status")
```






