---
title: "CH and HIV analysis"
author: "Christelle Colin-Leitzinger"
date: '`r Sys.Date()`'
output: 
  html_document:
    toc: true
    toc_float: true
    toc_collapsed: false
    theme: united
    highlight: pygments
    df_print: kable
editor_options: 
  chunk_output_type: console
---

<style type="text/css">

.figure {
    margin-top: 100px;
    margin-bottom: 100px;
}

table {
    margin-top: 10px;
    margin-bottom: 25px !important;
}

th, td { padding: 5px; }

</style>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      warning = FALSE,
                      message = FALSE,
                      cache = FALSE,
                      fig.align='center'
                      )
```

```{r library}
library(tidyverse)
library(gtsummary)
library(survival)
library(survminer)
theme_set(theme_classic())
theme_gtsummary_compact()
```

```{r}
CH_HIV_data <- 
  read_rds(paste0(here::here(), "/CH_HIV_data_full.rds")) %>% 
  group_by(MRN) %>% 
  mutate(n = n(), .before = MRN) %>% 
  ungroup() %>% 
  filter(baseline_sample == 1 |
           (n == 1 & baseline_sample == 2)) %>% 
  select(-n)
```

# Table 1 - Clinical in Overall Population
```{r table 1}
CH_HIV_data %>% 
  select(age_at_diagnosis_corrected, age_at_specimen,
         hiv_status, CH_status,
         starts_with("grim"), starts_with("tnsc"),
         starts_with("horvath"),
         starts_with("pheno_"),
         starts_with("dunedin"),
         new_race, new_ethnicity, sex, 
         new_tnm_stage, clinical_stage, 
         vital_status, tx_summary, new_tx_summary, 
         treatment_status,
         specimen_to_surgery_lag,
         specimen_to_treatment_lag, specimen_tx_lag,
         new_primary_site,
         DERIVED_TOBACCO_SMOKING_STATUS_DESC
         ) %>% 
  
  tbl_summary() %>% 
  bold_labels()
```

```{r plot clocks, fig.height=14, fig.width=12}
CH_HIV_data %>%
  select(CH_status,
         starts_with("grim"), starts_with("tnsc"),
         starts_with("horvath"),
         starts_with("pheno_"),
         starts_with("dunedin")) %>%
  filter(!is.na(CH_status)) %>%
  pivot_longer(cols = -CH_status) %>%
  ggplot(aes(x= CH_status, y= value, color= CH_status))+
  geom_boxplot()+
  scale_color_manual(values = c("blue", "red"))+
  facet_wrap(.~ name, scales = "free", ncol = 3)+
  theme(axis.title.y = element_blank())+
  stat_compare_means()

CH_HIV_data %>%
  select(hiv_status,
         starts_with("grim"), starts_with("tnsc"),
         starts_with("horvath"),
         starts_with("pheno_"),
         starts_with("dunedin")) %>%
  filter(!is.na(hiv_status)) %>%
  pivot_longer(cols = -hiv_status) %>%
  ggplot(aes(x= hiv_status, y= value, color= hiv_status))+
  geom_boxplot()+
  scale_color_manual(values = c("darkgrey", "purple"))+
  facet_wrap(.~ name, scales = "free", ncol = 3)+
  theme(axis.title.y = element_blank())+
  stat_compare_means()
```

# Table 2 - Clinical by CH
```{r table 2}
CH_HIV_data %>% 
  select(age_at_diagnosis_corrected, age_at_specimen,
         hiv_status, CH_status,
         starts_with("grim"), starts_with("tnsc"),
         starts_with("horvath"),
         starts_with("pheno_"),
         starts_with("dunedin"),
         new_race, new_ethnicity, sex, 
         new_tnm_stage, clinical_stage, 
         vital_status, tx_summary, new_tx_summary,
         treatment_status,
         specimen_to_surgery_lag,
         specimen_to_treatment_lag, specimen_tx_lag,
         new_primary_site,
         DERIVED_TOBACCO_SMOKING_STATUS_DESC
         ) %>% 
  tbl_summary(by = CH_status) %>% 
  bold_labels() %>% 
  add_p() %>% bold_p(t=.05) %>% add_overall()
```
New_primary_site Stratified by CH
```{r}
tbl1 <- CH_HIV_data %>% 
  filter(CH_status == "CH") %>% 
  select(hiv_status,
         new_primary_site
         ) %>% 
  tbl_summary(by = hiv_status) %>% 
  bold_labels() %>% 
  add_p() %>% bold_p(t=.05) %>% add_overall()
tbl2 <- CH_HIV_data %>% 
  filter(CH_status == "NO") %>% 
  select(hiv_status,
         new_primary_site
         ) %>% 
  tbl_summary(by = hiv_status) %>% 
  bold_labels() %>% 
  add_p() %>% bold_p(t=.05) %>% add_overall()
tbl_merge(list(tbl1, tbl2), tab_spanner = c("CH", "No CH"))
```
CH by Primary Site
```{r}
CH_HIV_data %>% 
  select(CH_status,
         new_primary_site
         ) %>% 
  tbl_summary(by = new_primary_site) %>% 
  bold_labels() %>% 
  add_p() %>% bold_p(t=.05) %>% add_overall()
```
CH _vs_ HIV Stratified by Primary Site
```{r}
tbl1 <- CH_HIV_data %>% 
  filter(new_primary_site == "Breast") %>% 
  select(CH_status,
         hiv_status
         ) %>% 
  tbl_summary(by = hiv_status) %>% 
  bold_labels() %>% 
  add_p() %>% bold_p(t=.05)
tbl2 <- CH_HIV_data %>% 
  filter(new_primary_site == "GI") %>% 
  select(CH_status,
         hiv_status
         ) %>% 
  tbl_summary(by = hiv_status) %>% 
  bold_labels() %>% 
  add_p() %>% bold_p(t=.05)
tbl3 <- CH_HIV_data %>% 
  filter(new_primary_site == "GU") %>% 
  select(CH_status,
         hiv_status
         ) %>% 
  tbl_summary(by = hiv_status) %>% 
  bold_labels() %>% 
  add_p() %>% bold_p(t=.05)
tbl4 <- CH_HIV_data %>% 
  filter(new_primary_site == "GYN") %>% 
  select(CH_status,
         hiv_status
         ) %>% 
  tbl_summary(by = hiv_status) %>% 
  bold_labels() %>% 
  add_p() %>% bold_p(t=.05)
tbl5 <- CH_HIV_data %>% 
  filter(new_primary_site == "Head and Neck") %>% 
  select(CH_status,
         hiv_status
         ) %>% 
  tbl_summary(by = hiv_status) %>% 
  bold_labels() %>% 
  add_p() %>% bold_p(t=.05)
tbl6 <- CH_HIV_data %>% 
  filter(new_primary_site == "Lung") %>% 
  select(CH_status,
         hiv_status
         ) %>% 
  tbl_summary(by = hiv_status) %>% 
  bold_labels() %>% 
  add_p() %>% bold_p(t=.05)
tbl7 <- CH_HIV_data %>% 
  filter(new_primary_site == "Skin") %>% 
  select(CH_status,
         hiv_status
         ) %>% 
  tbl_summary(by = hiv_status) %>% 
  bold_labels() %>% 
  add_p() %>% bold_p(t=.05)

tbl_merge(list(tbl1, tbl2, tbl3, tbl4, tbl5, tbl6, tbl7), 
          tab_spanner = c("Breast", "GI", "GU", "GYN", "Head and Neck", "Lung", "Skin"))
```

# Table 3 - Clinical by HIV
```{r}
CH_HIV_data %>% 
  select(age_at_diagnosis_corrected, age_at_specimen,
         hiv_status, CH_status,
         starts_with("grim"), starts_with("tnsc"),
         starts_with("horvath"),
         starts_with("pheno_"),
         starts_with("dunedin"),
         new_race, new_ethnicity, sex, 
         new_tnm_stage, clinical_stage, 
         vital_status, tx_summary, new_tx_summary, 
         treatment_status,
         specimen_to_surgery_lag,
         specimen_to_treatment_lag, specimen_tx_lag,
         new_primary_site,
         DERIVED_TOBACCO_SMOKING_STATUS_DESC
         ) %>% 
  tbl_summary(by = hiv_status) %>% 
  bold_labels() %>% 
  add_p() %>% bold_p(t=.05) %>% add_overall()
```

<!-- # Table 4 - stratified by CH and HIV -->


# Survival analysis
**Note : Survival time is calculated from date of treatment to date of death**  
<!-- <div class = "row"> -->
<!-- <div class = "col-md-6"> -->
<!-- ```{r Survival, fig.height = 7} -->
<!-- ggsurvplot(survfit(Surv(month_at_os, os_event) ~ CH_status,  -->
<!--                    data=CH_HIV_data), -->
<!--            title = "OS Analysis", -->
<!--            font.main = c(20, "bold", "black"), -->
<!--            font.x = c(18, "bold", "black"), -->
<!--            font.y = c(18, "bold", "black"), -->
<!--            font.legend = c(16, "black"), -->
<!--            font.tickslab = c(16, "bold", "black"), -->
<!--            size = 1.5, -->

<!--            xlab = "Time in months", -->
<!--            legend = "top", -->
<!--            legend.title = "", -->
<!--            pval = TRUE, -->
<!--            conf.int = FALSE, -->
<!--            # Censor -->
<!--            censor = TRUE -->
<!-- ) + guides(colour = guide_legend(ncol = 1)) -->
<!-- ``` -->
<!-- </div> -->

<!-- <div class = "col-md-6"> -->
<!-- ```{r, fig.height = 7} -->
<!-- ggsurvplot(survfit(Surv(month_at_os, os_event) ~ hiv_status,  -->
<!--                    data=CH_HIV_data), -->
<!--            title = "OS Analysis", -->
<!--            font.main = c(20, "bold", "black"), -->
<!--            font.x = c(18, "bold", "black"), -->
<!--            font.y = c(18, "bold", "black"), -->
<!--            font.legend = c(16, "black"), -->
<!--            font.tickslab = c(16, "bold", "black"), -->
<!--            size = 1.5, -->

<!--            xlab = "Time in months", -->
<!--            legend = "top", -->
<!--            legend.title = "", -->
<!--            pval = TRUE, -->
<!--            conf.int = FALSE, -->
<!--            # Censor -->
<!--            censor = TRUE -->
<!-- ) + guides(colour = guide_legend(ncol = 1)) -->
<!-- ``` -->
<!-- </div> -->
<!-- </div> -->

<!-- ## 1.MV Cox model w/ CH+HIV+age+race+eth+sex+clinicalstage -->
<!-- ```{r cox} -->
<!-- tbl1 <- CH_HIV_data %>%  -->
<!--   select(CH_status,  -->
<!--          age_at_diagnosis_corrected, -->
<!--          new_race, new_ethnicity, sex, hiv_status, clinical_stage) %>%  -->
<!--   tbl_uvregression(method = survival::coxph,  -->
<!--                    y = (Surv(time = CH_HIV_data$month_at_os,  -->
<!--                              event = CH_HIV_data$os_event)), -->
<!--                    exponentiate = TRUE) %>%  -->
<!--   bold_p(t = .05) %>% add_nevent(location = "level") %>% add_n(location = "level") %>% -->
<!--   bold_labels() %>% italicize_levels() -->
<!-- tbl2 <- coxph(Surv(time = CH_HIV_data$month_at_os,  -->
<!--              event = CH_HIV_data$os_event) ~  -->
<!--                CH_status + age_at_diagnosis_corrected + new_race + new_ethnicity + sex + hiv_status + clinical_stage,  -->
<!--              data =  CH_HIV_data) %>% -->
<!--   tbl_regression(exponentiate = TRUE) %>% bold_p(t = .05) -->

<!-- tbl_merge(list(tbl1, tbl2), tab_spanner = c("**Univariable**", "**Multivariable**")) -->
<!-- ``` -->

<!-- ## MV Cox model w/ CH+HIV -->
<!-- ```{r cox} -->
<!-- tbl1 <- CH_HIV_data %>%  -->
<!--   select(CH_status,  -->
<!--          hiv_status) %>%  -->
<!--   tbl_uvregression(method = survival::coxph,  -->
<!--                    y = (Surv(time = CH_HIV_data$month_at_os,  -->
<!--                              event = CH_HIV_data$os_event)), -->
<!--                    exponentiate = TRUE) %>%  -->
<!--   bold_p(t = .05) %>% add_nevent(location = "level") %>% add_n(location = "level") %>% -->
<!--   bold_labels() %>% italicize_levels() -->
<!-- tbl2 <- coxph(Surv(time = CH_HIV_data$month_at_os,  -->
<!--              event = CH_HIV_data$os_event) ~  -->
<!--                CH_status + hiv_status,  -->
<!--              data =  CH_HIV_data) %>% -->
<!--   tbl_regression(exponentiate = TRUE) %>% bold_p(t = .05) -->

<!-- tbl_merge(list(tbl1, tbl2), tab_spanner = c("**Univariable**", "**Multivariable**")) -->
<!-- ``` -->
<!-- Cox Proportion Hazard Assumption - No violation of the model -->
<!-- ```{r} -->
<!-- coxph(Surv(time = CH_HIV_data$month_at_os,  -->
<!--              event = CH_HIV_data$os_event) ~  -->
<!--                CH_status + age_at_diagnosis_corrected + new_race + new_ethnicity + sex + hiv_status + clinical_stage,  -->
<!--              data =  CH_HIV_data) %>%  -->
<!--   cox.zph() -->
<!-- ``` -->

## 1.MV Cox model w/ CH+HIV+age
```{r}
tbl1 <- CH_HIV_data %>%
  select(CH_status,
         age_at_diagnosis_corrected,
         hiv_status) %>%
  tbl_uvregression(method = survival::coxph,
                   y = (Surv(time = CH_HIV_data$month_at_os,
                             event = CH_HIV_data$os_event)),
                   exponentiate = TRUE) %>%
  bold_p(t = .05) %>% add_nevent(location = "level") %>% add_n(location = "level") %>%
  bold_labels() %>% italicize_levels()
tbl2 <- coxph(Surv(time = CH_HIV_data$month_at_os,
             event = CH_HIV_data$os_event) ~
               CH_status + age_at_diagnosis_corrected + hiv_status,
             data =  CH_HIV_data) %>%
  tbl_regression(exponentiate = TRUE) %>% bold_p(t = .05)

tbl_merge(list(tbl1, tbl2), tab_spanner = c("**Univariable**", "**Multivariable**"))
```
<br>
<br>

<!-- ## In HIV + population -->
<!-- ```{r} -->
<!-- HIV_pos <- CH_HIV_data %>%  -->
<!--   filter(hiv_status == "positive") -->
<!-- tbl1 <- HIV_pos %>%  -->
<!--   select(CH_status,  -->
<!--          age_at_diagnosis_corrected, -->
<!--          new_race, new_ethnicity, sex, clinical_stage) %>%  -->
<!--   tbl_uvregression(method = survival::coxph,  -->
<!--                    y = (Surv(time = HIV_pos$month_at_os,  -->
<!--                              event = HIV_pos$os_event)), -->
<!--                    exponentiate = TRUE) %>%  -->
<!--   bold_p(t = .05) %>% add_nevent(location = "level") %>% add_n(location = "level") %>% -->
<!--   bold_labels() %>% italicize_levels() -->
<!-- tbl2 <- coxph(Surv(time = HIV_pos$month_at_os,  -->
<!--              event = HIV_pos$os_event) ~  -->
<!--                CH_status + age_at_diagnosis_corrected + new_race + new_ethnicity + sex + clinical_stage,  -->
<!--              data =  HIV_pos) %>% -->
<!--   tbl_regression(exponentiate = TRUE) %>% bold_p(t = .05) -->

<!-- tbl_merge(list(tbl1, tbl2), tab_spanner = c("**Univariable**", "**Multivariable**")) -->
<!-- ``` -->

<!-- ## In HIV - population -->
<!-- ```{r} -->
<!-- HIV_neg <- CH_HIV_data %>%  -->
<!--   filter(hiv_status == "negative") -->
<!-- tbl1 <- HIV_neg %>%  -->
<!--   select(CH_status,  -->
<!--          age_at_diagnosis_corrected, -->
<!--          new_race, new_ethnicity, sex, clinical_stage) %>%  -->
<!--   tbl_uvregression(method = survival::coxph,  -->
<!--                    y = (Surv(time = HIV_neg$month_at_os,  -->
<!--                              event = HIV_neg$os_event)), -->
<!--                    exponentiate = TRUE) %>%  -->
<!--   bold_p(t = .05) %>% add_nevent(location = "level") %>% add_n(location = "level") %>% -->
<!--   bold_labels() %>% italicize_levels() -->
<!-- tbl2 <- coxph(Surv(time = HIV_neg$month_at_os,  -->
<!--              event = HIV_neg$os_event) ~  -->
<!--                CH_status + age_at_diagnosis_corrected + new_race + new_ethnicity + sex + clinical_stage,  -->
<!--              data =  HIV_neg) %>% -->
<!--   tbl_regression(exponentiate = TRUE) %>% bold_p(t = .05) -->

<!-- tbl_merge(list(tbl1, tbl2), tab_spanner = c("**Univariable**", "**Multivariable**")) -->
<!-- ``` -->

<!-- ## In No CH population -->
<!-- ```{r} -->
<!-- No_CH <- CH_HIV_data %>%  -->
<!--   filter(CH_status == "NO") -->
<!-- tbl1 <- No_CH %>%  -->
<!--   select(hiv_status,  -->
<!--          age_at_diagnosis_corrected, -->
<!--          new_race, new_ethnicity, sex, clinical_stage) %>%  -->
<!--   tbl_uvregression(method = survival::coxph,  -->
<!--                    y = (Surv(time = No_CH$month_at_os,  -->
<!--                              event = No_CH$os_event)), -->
<!--                    exponentiate = TRUE) %>%  -->
<!--   bold_p(t = .05) %>% add_nevent(location = "level") %>% add_n(location = "level") %>% -->
<!--   bold_labels() %>% italicize_levels() -->
<!-- tbl2 <- coxph(Surv(time = No_CH$month_at_os,  -->
<!--              event = No_CH$os_event) ~  -->
<!--                hiv_status + age_at_diagnosis_corrected + new_race + new_ethnicity + sex + clinical_stage,  -->
<!--              data =  No_CH) %>% -->
<!--   tbl_regression(exponentiate = TRUE) %>% bold_p(t = .05) -->

<!-- tbl_merge(list(tbl1, tbl2), tab_spanner = c("**Univariable**", "**Multivariable**")) -->
<!-- ``` -->

<!-- ## In CH population -->
<!-- ```{r} -->
<!-- CH_data <- CH_HIV_data %>%  -->
<!--   filter(CH_status == "CH") -->
<!-- tbl1 <- CH_data %>%  -->
<!--   select(hiv_status,  -->
<!--          age_at_diagnosis_corrected, -->
<!--          new_race, new_ethnicity, sex, clinical_stage) %>%  -->
<!--   tbl_uvregression(method = survival::coxph,  -->
<!--                    y = (Surv(time = CH_data$month_at_os,  -->
<!--                              event = CH_data$os_event)), -->
<!--                    exponentiate = TRUE) %>%  -->
<!--   bold_p(t = .05) %>% add_nevent(location = "level") %>% add_n(location = "level") %>% -->
<!--   bold_labels() %>% italicize_levels() -->
<!-- tbl2 <- coxph(Surv(time = CH_data$month_at_os,  -->
<!--              event = CH_data$os_event) ~  -->
<!--                hiv_status + age_at_diagnosis_corrected + new_race + new_ethnicity + sex + clinical_stage,  -->
<!--              data =  CH_data) %>% -->
<!--   tbl_regression(exponentiate = TRUE) %>% bold_p(t = .05) -->

<!-- tbl_merge(list(tbl1, tbl2), tab_spanner = c("**Univariable**", "**Multivariable**")) -->
<!-- ``` -->

## 2.Survival curve
```{r}
ggsurvplot(survfit(Surv(month_at_os, os_event) ~ CH_status, 
                   data=CH_HIV_data),
           title = "OS Analysis",
           font.main = c(20, "bold", "black"),
           font.x = c(18, "bold", "black"),
           font.y = c(18, "bold", "black"),
           font.legend = c(16, "black"),
           font.tickslab = c(16, "bold", "black"),
           size = 1.5,

           xlab = "Time in months",
           legend = "top",
           legend.title = "CH status",
           legend.labs = c("No", "CH"),
           palette = c("blue", "red"),
           pval = TRUE,
           conf.int = FALSE,
           # Censor
           censor = TRUE
) + guides(colour = guide_legend(ncol = 1))
```

```{r}
ggsurvplot(survfit(Surv(month_at_os, os_event) ~ hiv_status, 
                   data=CH_HIV_data),
           title = "OS Analysis",
           font.main = c(20, "bold", "black"),
           font.x = c(18, "bold", "black"),
           font.y = c(18, "bold", "black"),
           font.legend = c(16, "black"),
           font.tickslab = c(16, "bold", "black"),
           size = 1.5,

           xlab = "Time in months",
           legend = "top",
           legend.title = "HIV status",
           legend.labs = c("negative", "positive"),
           palette = c("darkgrey", "purple"),
           pval = TRUE,
           conf.int = FALSE,
           # Censor
           censor = TRUE
) + guides(colour = guide_legend(ncol = 1))
```

## 3.Survival by HIV/CH
```{r}
ggsurvplot(survfit(Surv(month_at_os, os_event) ~ hiv_status + CH_status,
                   data=CH_HIV_data),
           title = "OS Analysis",
           font.main = c(20, "bold", "black"),
           font.x = c(18, "bold", "black"),
           font.y = c(18, "bold", "black"),
           font.legend = c(16, "black"),
           font.tickslab = c(16, "bold", "black"),
           size = 1.5,

           xlab = "Time in months",
           legend = "top",
           legend.title = "",
           color = "hiv_status",
           palette = c("darkgrey", "purple"),
           linetype = "CH_status",
           pval = TRUE,
           conf.int = FALSE,
           # Censor
           censor = TRUE
) + guides(colour = guide_legend(ncol = 1),
           linetype = guide_legend(ncol = 1))
```

# Logistic regression
```{r}
# glm(CH_status ~
#       hiv_status + age_at_diagnosis_corrected + new_race + sex + treatment_status,
#     family = "binomial",
#     data = CH_HIV_data) %>%
#   tbl_regression(exponentiate = TRUE,
#                  intercept = TRUE) %>%
#   modify_spanning_header(everything() ~ "CH status")

glm(CH_status ~
      hiv_status + age_at_diagnosis_corrected + grim_age_accel,
    family = "binomial",
    data = CH_HIV_data) %>%
  tbl_regression(exponentiate = TRUE,
                 intercept = TRUE) %>%
  modify_spanning_header(everything() ~ "CH status")

glm(CH_status ~
      hiv_status + age_at_diagnosis_corrected + tnsc_accel,
    family = "binomial",
    data = CH_HIV_data) %>%
  tbl_regression(exponentiate = TRUE,
                 intercept = TRUE) %>%
  modify_spanning_header(everything() ~ "CH status")

glm(CH_status ~
      hiv_status + age_at_diagnosis_corrected + pheno_age_accel,
    family = "binomial",
    data = CH_HIV_data) %>%
  tbl_regression(exponentiate = TRUE,
                 intercept = TRUE) %>%
  modify_spanning_header(everything() ~ "CH status")

glm(CH_status ~
      hiv_status + age_at_diagnosis_corrected + dunedin_pace,
    family = "binomial",
    data = CH_HIV_data) %>%
  tbl_regression(exponentiate = TRUE,
                 intercept = TRUE) %>%
  modify_spanning_header(everything() ~ "CH status")

# glm(CH_status ~
#       hiv_status + age_at_diagnosis_corrected + grim_age_accel + tnsc_accel + pheno_age_accel + dunedin_pace,
#     family = "binomial",
#     data = CH_HIV_data) %>%
#   tbl_regression(exponentiate = TRUE,
#                  intercept = TRUE) %>%
#   modify_spanning_header(everything() ~ "CH status")

glm(CH_status ~
      hiv_status + horvath18,
    family = "binomial",
    data = CH_HIV_data) %>%
  tbl_regression(exponentiate = TRUE,
                 intercept = TRUE) %>%
  modify_spanning_header(everything() ~ "CH status")

glm(CH_status ~
      hiv_status + horvath13,
    family = "binomial",
    data = CH_HIV_data) %>%
  tbl_regression(exponentiate = TRUE,
                 intercept = TRUE) %>%
  modify_spanning_header(everything() ~ "CH status")

# glm(CH_status ~
#       hiv_status + horvath18 + horvath13,
#     family = "binomial",
#     data = CH_HIV_data) %>%
#   tbl_regression(exponentiate = TRUE,
#                  intercept = TRUE) %>%
#   modify_spanning_header(everything() ~ "CH status")
```

```{r}
lung_data <- CH_HIV_data %>% 
   filter(new_primary_site == "Lung")
print("table smoking vs CH")
table(lung_data$DERIVED_TOBACCO_SMOKING_STATUS_DESC, lung_data$CH_status)

print("table smoking vs HIV")
table(lung_data$DERIVED_TOBACCO_SMOKING_STATUS_DESC, lung_data$hiv_status)

glm(CH_status ~
      hiv_status + DERIVED_TOBACCO_SMOKING_STATUS_DESC,
    family = "binomial",
    data = lung_data) %>%
  tbl_regression(exponentiate = TRUE,
                 intercept = TRUE) %>%
  modify_spanning_header(everything() ~ "CH status")
```

# Logistic regression model: CH outcome, model to include HIV, race, sex, and smoking status only.
```{r}
glm(CH_status ~
      hiv_status + new_race + sex + DERIVED_TOBACCO_SMOKING_STATUS_DESC,
    family = "binomial",
    data = CH_HIV_data) %>%
  tbl_regression(exponentiate = TRUE,
                 intercept = TRUE) %>%
  modify_spanning_header(everything() ~ "CH status")
```

# Cox PH regression model: Survival outcome, model to include CH, race, sex, and smoking status only. Let’s do both overall and only among HIV+ patients.
```{r}
tbl1 <- CH_HIV_data %>%
  select(CH_status,
         new_race, sex,
         DERIVED_TOBACCO_SMOKING_STATUS_DESC) %>%
  tbl_uvregression(method = survival::coxph,
                   y = (Surv(time = CH_HIV_data$month_at_os,
                             event = CH_HIV_data$os_event)),
                   exponentiate = TRUE) %>%
  bold_p(t = .05) %>% add_nevent(location = "level") %>% add_n(location = "level") %>%
  bold_labels() %>% italicize_levels()
tbl2 <- coxph(Surv(time = CH_HIV_data$month_at_os,
             event = CH_HIV_data$os_event) ~
               CH_status + new_race + sex + DERIVED_TOBACCO_SMOKING_STATUS_DESC,
             data =  CH_HIV_data) %>%
  tbl_regression(exponentiate = TRUE) %>% bold_p(t = .05) %>%
  modify_caption("**Overall population**")

tbl_merge(list(tbl1, tbl2), tab_spanner = c("**Univariable**", "**Multivariable**"))
```

```{r}
HIV_pos <- CH_HIV_data %>%
  filter(hiv_status == "positive")

tbl1 <- HIV_pos %>%
  select(CH_status,
         new_race, sex,
         DERIVED_TOBACCO_SMOKING_STATUS_DESC) %>%
  tbl_uvregression(method = survival::coxph,
                   y = (Surv(time = HIV_pos$month_at_os,
                             event = HIV_pos$os_event)),
                   exponentiate = TRUE) %>%
  bold_p(t = .05) %>% add_nevent(location = "level") %>% add_n(location = "level") %>%
  bold_labels() %>% italicize_levels() %>% 
  modify_header(label = "**Cox PH HIV+**")
tbl2 <- coxph(Surv(time = HIV_pos$month_at_os,
             event = HIV_pos$os_event) ~
               CH_status + new_race + sex + DERIVED_TOBACCO_SMOKING_STATUS_DESC,
             data =  HIV_pos) %>%
  tbl_regression(exponentiate = TRUE) %>% bold_p(t = .05) %>%
  modify_caption("**HIV+**")

tbl_merge(list(tbl1, tbl2), tab_spanner = c("**Univariable**", "**Multivariable**"))
```

